---
title: "Capstone Graphs"
output: html_document
date: "2023-03-27"
---
Reading in the data for all three Species and loading in libraries
```{r}
archcow=read.csv('ArchCow.csv')
archcow=archcow[,1:6]
View(archcow)
moderncow=read.csv('ModernCow.csv')
View(moderncow)
archelephant=read.csv('ArchElephant.csv')
View(archelephant)
library(ggplot2)
```

Archeological Cow Carbon Vs Machine Type Analysis 
```{r}
ac13cmachinelm=lm(X13C~as.factor(Machine),data=archcow)
summary(ac13cmachinelm)
anova(ac13)
acCarbonMachineTest=t.test(archcow[archcow$Machine=='p',2],archcow[archcow$Machine=='GasBench',2],var.equal=F)
acCarbonMachineTest
ggplot()+geom_density(aes(x=archcow[archcow$Machine=='p',2]),color='red')+geom_density(aes(x=archcow[archcow$Machine=='GasBench',2]),color='blue')
```
Archeological Cow Oxygen Vs Machine Type Analysis
```{r}
ac18Omachinelm=lm(X18O~as.factor(Machine),data=archcow)
summary(ac18Omachinelm)
acOxygenMachineTest=t.test(archcow[archcow$Machine=='p',3],archcow[archcow$Machine=='GasBench',3],var.equal=F)
acOxygenMachineTest
ggplot()+geom_density(aes(x=archcow[archcow$Machine=='p',3]),color='red')+geom_density(aes(x=archcow[archcow$Machine=='GasBench',3]),color='blue')
```
Modern Cow Carbon Vs Machine Type Analysis
```{r}
mc13cmachinelm=lm(X13C~as.factor(Machine),data=moderncow)
summary(mc13cmachinelm)
mcCarbonMachineTest=t.test(moderncow[moderncow$Machine=='p',2],moderncow[moderncow$Machine=='GasBench',2],var.equal=F)
mcCarbonMachineTest
ggplot()+geom_density(aes(x=moderncow[moderncow$Machine=='p',2]),color='red')+geom_density(aes(x=moderncow[moderncow$Machine=='GasBench',2]),color='blue')
```
Modern Cow Oxygen Vs Machine Type Analysis
```{r}
mc18Omachinelm=lm(X18O~as.factor(Machine),data=moderncow)
summary(mc18Omachinelm)
mcOxygenMachineTest=t.test(moderncow[moderncow$Machine=='p',3],moderncow[moderncow$Machine=='GasBench',3],var.equal=F)
mcOxygenMachineTest
ggplot()+geom_density(aes(x=moderncow[moderncow$Machine=='p',3]),color='red')+geom_density(aes(x=moderncow[moderncow$Machine=='GasBench',3]),color='blue')
```
Archeologcial Elephant Carbon Vs Machine Type Analysis
```{r}
ae13cmachinelm=lm(X13C~as.factor(Machine),data=archelephant)
summary(ae13cmachinelm)
aeCarbonMachineTest=t.test(archelephant[archelephant$Machine=='p',2],archelephant[archelephant$Machine=='GasBench',2],var.equal=F)
aeCarbonMachineTest
ggplot()+geom_density(aes(x=archelephant[archelephant$Machine=='p',2]),color='red')+geom_density(aes(x=archelephant[archelephant$Machine=='GasBench',2]),color='blue')
```
Archeological Elephant Oxygen Vs Machine Type Analysis
```{r}
ae18Omachinelm=lm(X18O~as.factor(Machine),data=archelephant)
summary(ae18Omachinelm)
aeOxygenMachineTest=t.test(archelephant[archelephant$Machine=='p',3],archelephant[archelephant$Machine=='GasBench',3],var.equal=F)
aeOxygenMachineTest
ggplot()+geom_density(aes(x=archelephant[archelephant$Machine=='p',3]),color='red')+geom_density(aes(x=archelephant[archelephant$Machine=='GasBench',3]),color='blue')
```

Scatter plot for each species and each isotope by year
```{r}
ggpabsc13Xy=ggplot(archcow,aes(y=X13C,x=Year))+geom_point()+labs(title='Scatterplot for Archeological Cow',y='Delta Carbon 13 ratio',x='Year \n Figure 13')
ggpmbsc13Xy=ggplot(moderncow,aes(y=X13C,x=Year))+geom_point()+labs(title='Scatterplot for Modern Cow',y='Delta Carbon 13 ratio',x='Year \n Figure 14')
ggpaesc13Xy=ggplot(archelephant,aes(y=X13C,x=Year))+geom_point()+labs(title='Scatterplot for Archeological Elephant',y='Delta Carbon 13 ratio',x='Year \n Figure 15')
grid.arrange(ggpabsc13Xy,ggpmbsc13Xy,ggpaesc13Xy,nrow=1)
 ggpabso18Xy=ggplot(archcow,aes(y=X18O,x=Year))+geom_point()+labs(title='Scatterplot for Archeological Cow',y='Delta Oxygen 18 ratio',x='Year \n Figure 10')
ggpmbso18Xy=ggplot(moderncow,aes(y=X18O,x=Year))+geom_point()+labs(title='Scatterplot for Modern Cow',y='Delta Oxygen 18 ratio',x='Year \n Figure 11')
ggpaeso18Xy=ggplot(archelephant,aes(y=X18O,x=Year))+geom_point()+labs(title='Scatterplot for Archeological Elephant',y='Delta Oxygen 18 ratio',x='Year \n Figure 12')
grid.arrange(ggpabso18Xy,ggpmbso18Xy,ggpaeso18Xy,nrow=1)
```


```{r}
```




