---
title: "Final Project"
output:
  word_document: default
  html_document: default
date: "2022-12-02"
---
```{r}
#############################
## Final Project Code ##
#############################

## Install and load packages
library(rjags)
library(MCMCvis)
library(dplyr)
## Load cat student data
cat=read.table("cats.txt",header=T)
cat$Breed=recode_factor(cat$Breed,A=1,B=1,C=1,D=0,E=0)
attach(cat)
n <- 40  # sample size
p <- dim(cat)[2]-1  # number of regression coefficients

#####################################################
## Linear regression model ##
#####################################################
## (1) Create data list
dataList1 <- list(
        "n" = n,
        "DeltaPCV" = DeltaPCV,
        "Dose" = Dose,
        "Type" = Type,
        "Breed" = Breed)

## (2) Specify list of parameter(s) to be monitored
parameters1 <- c("beta","sig2")

## (3) Specify initial values for parameter(s) in Metropolis-Hastings algorithm
initsValues1 <- list(
        "beta" = rep(0,p+1),
        "inv_sig2" = 1
)

## (4) Specify parameters for running Metropolis-Hastings algorithm
adaptSteps1 <- 10000              # number of steps to "tune" the samplers
burnInSteps1 <- 20000             # number of steps to "burn-in" the samplers
nChains1 <- 3                    # number of chains to run
numSavedSteps1 <- 50000          # total number of steps in chains to save
thinSteps1 <- 10                  # number of steps to "thin" (1 = keep every step)
nIter <- ceiling((numSavedSteps1*thinSteps1)/nChains1) 	# steps per chain

## (5) Create, initialize, and adapt the model
# This will require you to create a separate .txt file which specifies
# the model
jagsModelF <- jags.model("Final_Project_Model.txt", 
                         data = dataList1, 
                         inits = initsValues1, 
                         n.chains = nChains1, 
                         n.adapt = adaptSteps1)
jagsModelF
## (6) Burn-in the algorithm
if(burnInSteps1>0){
        cat( "Burning in the MCMC chain...\n")
        update(jagsModelF, n.iter = burnInSteps1)
}

## (7) Run MCMC algorithm
cat("Sampling final MCMC chain...\n" )
codaSamplesF <- coda.samples(jagsModelF, 
                             variable.names = parameters1, 
                             n.iter = nIter, 
                             thin = thinSteps1)

## (8) Diagnose convergence and plot posterior densities
par(ask=T)
plot(codaSamplesF)

## (9) Calculate numerical summaries for the posterior samples
summary(codaSamplesF)

## (10) Retrieve posterior samples for later use
mcmcChainF <- as.matrix(codaSamplesF)

## (11) Other plots
# Boxplot of posteriors for beta
boxplot(mcmcChainF[,1:5], main = "Posterior samples of beta",
        names=c("Intercept","Dose","Type","Breed","Sig2"))
abline(h = 0, lty = 2, col="red")

# 50% (thin line) and 95% (thick line) credible interval "caterpillar" plots
# dot is the posterior median
MCMCplot(codaSamplesF, params = "beta",
         main = "Posterior CIs for beta", ref = NULL)
```


```{r}

## Install and load packages
library(rjags)
library(MCMCvis)
library(dplyr)
## Load cat student data
cat=read.table("cats.txt",header=T)
cat$Breed=recode_factor(cat$Breed,A=1,B=1,C=1,D=0,E=0)
attach(cat)
n <- 40  # sample size
p <- dim(cat)[2]-1  # number of regression coefficients
# taking out School and Y but add one for intercept term

#####################################################
## Linear regression model ##
#####################################################
## (1) Create data list
dataList1 <- list(
  "n" = n,
  "DeltaPCV" = DeltaPCV,
  "Dose" = Dose,
  "Type" = Type,
  "Breed" = Breed)

## (2) Specify list of parameter(s) to be monitored
parameters1 <- c("beta","sig2")

## (3) Specify initial values for parameter(s) in Metropolis-Hastings algorithm
initsValues1F <- list(
  "beta" = rep(0,p+3),
  "inv_sig2" = .01
)

## (4) Specify parameters for running Metropolis-Hastings algorithm
adaptSteps1 <- 10000              # number of steps to "tune" the samplers
burnInSteps1 <- 20000             # number of steps to "burn-in" the samplers
nChains1 <- 3                    # number of chains to run
numSavedSteps1 <- 50000          # total number of steps in chains to save
thinSteps1 <- 10                  # number of steps to "thin" (1 = keep every step)
nIter <- ceiling((numSavedSteps1*thinSteps1)/nChains1) 	# steps per chain

## (5) Create, initialize, and adapt the model
# This will require you to create a separate .txt file which specifies
# the model
jagsModelF1 <- jags.model("Final_Project_Model_1.txt", 
                         data = dataList1, 
                         inits = initsValues1F, 
                         n.chains = nChains1, 
                         n.adapt = adaptSteps1)
jagsModelF1
## (6) Burn-in the algorithm
if(burnInSteps1>0){
  cat( "Burning in the MCMC chain...\n")
  update(jagsModelF1, n.iter = burnInSteps1)
}

## (7) Run MCMC algorithm
cat("Sampling final MCMC chain...\n" )
codaSamplesF1 <- coda.samples(jagsModelF1, 
                             variable.names = parameters1, 
                             n.iter = nIter, 
                             thin = thinSteps1)

## (8) Diagnose convergence and plot posterior densities
par(mar=c(1,1,1,1))
plot(codaSamplesF1)

## (9) Calculate numerical summaries for the posterior samples
summary(codaSamplesF1)

## (10) Retrieve posterior samples for later use
mcmcChainF1 <- as.matrix(codaSamplesF1)

## (11) Other plots
# Boxplot of posteriors for beta
boxplot(mcmcChainF1[,1:7], main = "Posterior samples of beta",
        names=c("Intercept","Dose","Type","Breed","Breed*Type","Breed*Dose","Sig2"))
abline(h = 0, lty = 2, col="red")

# 50% (thin line) and 95% (thick line) credible interval "caterpillar" plots
# dot is the posterior median
MCMCplot(codaSamplesF1, params = "beta",
         main = "Posterior CIs for beta", ref = NULL)

